Large squeezing in multicomponent trapped Bose-Einstein condensates

### 2012-11-24 large_number_squeezing_with_atoms.tex
Fixed multimode discussion

The first problem is present even if the multiple spatial modes are uncoupled.  The two-mode model shows that the parameters required to obtain best squeezing are dependent on the strength of the nonlinearity.  This nonlinearity is proportional to the product of the mode shape and the number of atoms in that mode.  The optimum hold time $\tau_{\mathrm{hold}}$ depends on this product, as does the recombination phase $\phi$.  Different modes in a multimode environment will typically varying products of nonlinearity and therefore squeeze at different rates, with each mode having a different optimum $\tau_{\mathrm{hold}}$ and optimum recombination phase $\phi$.  This means that best squeezing will be achieved by choosing values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across the range of densities, result in an overall lowest possible total number variance.

The second problem is one of the number-dependent dynamics of the spatial modes, which is observable even in a semiclassical simulation.  Due to the nonlinear term in the Hamiltonian, the local phase evolution of the atomic field is density dependent. If we write the order parameter of the atomic field as $\psi_i = \sqrt{\rho({\mathbf{r}})} \exp (i\varphi)$, where $j$ denotes one of the two states of our atomic field, the local phase evolution is given by
\begin{equation}
\frac{d \varphi({\mathbf{r}})}{dt} = \frac{U_{ii} \rho({\mathbf{r}})} {2 \hbar}.
\end{equation}
Consequently the mode shape of the atomic field changes depending on the nonlinearity-density product, which means that unless $U_{aa}|\psi_a({\mathbf{r}})|^2 = U_{bb}|\psi_b({\mathbf{r}})|^2$ for all ${\mathbf{r}}$, the mode shapes of will not overlap perfectly during the final coupling pulse. This degraded mode-matching will result in lower efficiency in converting the quadrature squeezing to number squeezing. We show an example of this effect in Figure~\ref{figXXX}.

>>>>

The first problem is present even if the spatial modes are uncoupled.  The two-mode model shows that the parameters required to obtain best squeezing are dependent on the strength of the nonlinearity.  This nonlinearity is a function of the mode shape and the number of atoms in that mode.  The optimum hold time $\tau_{\mathrm{hold}}$ depends on this product, as does the recombination phase $\phi$.  Different modes in a multimode environment will typically have varying products of nonlinearity and therefore squeeze at different rates, with each mode having a different optimum $\tau_{\mathrm{hold}}$ and optimum recombination phase $\phi$.  This means that the best squeezing will be achieved by choosing values of $\tau_{\mathrm{hold}}$ and $\phi$ that, averaged across the range of densities, result in an overall lowest possible total number variance.

The second problem is due to the number-dependent dynamics of the spatial modes, which is observable even in a semiclassical simulation.  Due to the nonlinear term in the Hamiltonian, the local phase evolution of the atomic field is density dependent.  Consequently the mode shape of the atomic field changes depending on the nonlinearity-density product, which means that unless $U_{aa}|\psi_a({\mathbf{r}})|^2 = U_{bb}|\psi_b({\mathbf{r}})|^2$ for all ${\mathbf{r}}$, the mode shapes will not overlap perfectly during the final coupling pulse. This degraded mode-matching will result in lower efficiency in converting the quadrature squeezing to number squeezing. We show an example of this effect in Figure~\ref{figXXX}.
